---
title: "DMA Assignment 1"
author: "Adeniyi Richard Michael-Adenuga"
date: "5/2/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(readr)
library(readxl)
library(kableExtra)
library(lubridate)
library(rfm)
library(ggfortify)
```

Read in the data:
```{r}
orders <- read_csv("~/Desktop/BusinessAnalytics /Term 3/Digital Marketing/Data/orders.csv", col_names = TRUE)

contacts <- read_csv("~/Desktop/BusinessAnalytics /Term 3/Digital Marketing/Data/contacts.csv", col_names = TRUE)

lines<-read_csv("~/Desktop/BusinessAnalytics /Term 3/Digital Marketing/Data/lines.csv", col_names = TRUE)

summary<-read_csv("~/Desktop/BusinessAnalytics /Term 3/Digital Marketing/Data/summary.csv", col_names = TRUE)

#Format the orderdate in the orders table and contactdate in the contacts table

orders$orderdate <- as.Date(orders$orderdate, format = "%Y - %m - %d")

lines$orderdate <- as.Date(lines$orderdate, format = "%Y - %m - %d")

contacts$contactdate <- as.Date(contacts$contactdate, format = "%Y - %m - %d")

```

# Q2ai Response Rate

## Distribution of days between consecutive catalog mails or emails

### Email
```{r}
email_contact<- contacts %>% filter(contacttype=="E")  #Extract Emails
catalog_contact <- contacts %>% filter(contacttype =="C") #Extract Catalog 

email_contact <- email_contact %>% mutate(year = format(contactdate, "%Y"))

email_frequency <- email_contact %>% group_by(cust_id,year) %>% arrange(contactdate) %>% summarise(days=mean((contactdate -lag(contactdate)),na.rm = TRUE))

email_frequency$days <- round(as.numeric(email_frequency$days),2)

email_frequency %>% ggplot(aes(x=days, color=year) ) + geom_density() + xlim(min=0,max=50) +theme_bw()

```

### Catalog
```{r}
catalog_contact <- catalog_contact %>% mutate(year = format(contactdate, "%Y"))

catalog_frequency <- catalog_contact %>% group_by(cust_id,year) %>% arrange(contactdate) %>% summarise(days=mean((contactdate -lag(contactdate)),na.rm = TRUE))

catalog_frequency$days <- round(as.numeric(catalog_frequency$days),2)

catalog_frequency %>% ggplot(aes(x=days, color=year) ) + geom_density() + xlim(min=0,max=50) +theme_bw()
```

### Response Rates

In order to remove the need to consider interactions between emails and catalog mailing, we focus on cases where we can isolate emails and also catalog mails. 

Since we are told that phone and mail are the only way to place catalog orders, we combine this with the knowledge of those customers to whom catalogs were sent to get an idea of catalog only.

```{r}
phone_mail_orders <- orders %>% filter(ordermethod=="P" | ordermethod=="M")
catalog_only_CIDs<- intersect(phone_mail_orders$cust_id,catalog_contact$cust_id)
catalog_only <- catalog_contact %>% filter(cust_id %in% catalog_only_CIDs )
```

#### Catalog Only:
```{r}
combined_catalog <- inner_join(phone_mail_orders, catalog_only, by= "cust_id") 
combined_catalog$diff <-combined_catalog$orderdate - combined_catalog$contactdate
required_catalog_extract<-combined_catalog %>% filter (between(diff,0,22))
```
Calculate responsiveness:
```{r}
catalog_table<- required_catalog_extract %>% group_by(year) %>% summarise(count=n())
catalog_table$catalogs_sent<- c(194650, 217354, 211790)
catalog_table$responsiveness <- (catalog_table$count/catalog_table$catalogs_sent)*100
catalog_table <- as.data.frame(catalog_table)
```

#### Email Only:
Find intersection between email CIDs and catalog CIDs. Combine the intersection with catalog only and then take the remaining as email.
```{r}
both_targeted <- intersect(email_contact$cust_id,catalog_only)
both_CIDs <- rbind(catalog_only_CIDs,both_targeted)
email_only <- email_contact %>% subset(!(cust_id %in% both_CIDs))

combined <- inner_join(orders, email_only, by= "cust_id")
combined$diff <-combined$orderdate - combined$contactdate
required_email_extract<-combined %>% filter (between(diff,0,4))
```
Calculate responsiveness:
```{r}
email_table<- required_email_extract %>% group_by(year) %>% summarise(count=n())
email_table$emails_sent<- c(343379, 464134, 534209)
email_table$responsiveness <- (email_table$count/email_table$emails_sent)*100
email_table <- as.data.frame(email_table)
```

Plot both:
```{r}
ggplot() + geom_line(data=email_table, aes(x=year, y=responsiveness, col="blue", group=1)) + geom_line(data=catalog_table, aes(x=year, y=responsiveness, col="red", group=1)) + scale_color_identity(name= "Channel",labels=c("Email", "Catalog Mail"),guide = "legend") + xlab(" Year") + ylab ("Response Rate (%)") + theme_bw()
```

# Q2b RFM

## RFM
```{r}
orders_new <- inner_join(orders, lines, by= "ordernum")
```

```{r}
current_date <- max(orders_new$orderdate.x)

orders_new <- orders_new  %>% mutate(year = format(orderdate.x, "%Y"))

orders_new2 <- orders_new %>% filter(year== 2007 | year==2008)

RFM <- orders_new2 %>% group_by(cust_id.x) %>% summarize(Recency = min(current_date-orderdate.x), Frequency= n(), Avg_spend=mean(linedollars, na.rm = TRUE))
```

Add the ranks
```{r}
RFM$R <- ntile(-RFM$Recency,5)
RFM$F <- ntile(RFM$Frequency,5)
RFM$M <- ntile(RFM$Avg_spend,5)
```
Combine for RFM score
```{r}
RFM <- RFM %>% mutate(RFM = paste0(R, F, M))
```

```{r}
RFM %>% ggplot() + geom_density(aes(x=Recency)) +xlim(min=0, max=400) + theme_bw()
RFM %>% ggplot() + geom_density(aes(x=Frequency)) + xlim(min=0, max=30)+theme_bw()
RFM %>% ggplot() + geom_density(aes(x=Avg_spend))+ xlim(min=0, max=250)+ xlab("Average Spend")+theme_bw()
```

## Put into RFM Segments

Segments as definied in the body of the report.

```{r}
RFM$Index<- 1:nrow(RFM)
```

champ
```{r}
champ<-RFM %>% filter(between(R,4,5) & between(M,4,5) & between(F,4,5))
RFM$segment<-NA

for(i in champ$Index){
  RFM[i,10] <- "Champions"
}
```

Loyal customers
```{r}
LC<-RFM %>% filter(between(R,2,5) & between(M,3,5) & between(F,3,5))

for(i in LC$Index){
  RFM[i,10] <- "Loyal Customers"
}
```

Potential Loyalist
```{r}
PL<-RFM %>% filter(between(R,3,5) & between(F,1,3) & between(M,1,3))

for(i in PL$Index){
  RFM[i,10] <- "Potential Loyalist"
}
```

Frugal Spenders
```{r}
FS<-RFM %>% filter(between(R,3,5) & between(F,3,5) & between(M,1,2))

for(i in FS$Index){
  RFM[i,10] <- "Frugal Spenders"
}
```
New Customers
```{r}
NC<-RFM %>% filter(between(R,4,5) & between(F,0,1) & between(M,0,1))

for(i in NC$Index){
  RFM[i,10] <- "New Customer"
}
```
Promising
```{r}
P<-RFM %>% filter(between(R,3,4) & between(F,0,1) & between(M,0,1))

for(i in P$Index){
  RFM[i,10] <- "Promising"
}
```
Need Attention
```{r}
N_A<-RFM %>% filter(between(R,2,3) & between(F,2,3) & between(M,2,3))

for(i in N_A$Index){
  RFM[i,10] <- "Need Attention"
}
```
About to sleep
```{r}
AS<-RFM %>% filter(between(R,2,3) & between(F,0,2) & between(M,0,2))

for(i in AS$Index){
  RFM[i,10] <- "About To Sleep"
}
```
At Risk
```{r}
AR<-RFM %>% filter(between(R,0,2) & between(F,2,5) & between(M,2,5))

for(i in AR$Index){
  RFM[i,10] <- "At Risk"
}
```
Can't Lose Them
```{r}
CLT<-RFM %>% filter(between(R,0,1) & between(F,4,5) & between(M,4,5))

for(i in CLT$Index){
  RFM[i,10] <- "Can't Lose Them"
}
```
Hibernating
```{r}
H<-RFM %>% filter(between(R,1,2) & between(F,1,2) & between(M,1,2))

for(i in H$Index){
  RFM[i,10] <- "Hibernating"
}
```
Lost
```{r}
L<-RFM %>% filter(between(R,0,2) & between(F,0,2) & between(M,0,2))

for(i in L$Index){
  RFM[i,10] <- "Lost"
}
```

New Big Spender
```{r}
NBS<-RFM %>% filter(between(R,3,5) & between(F,0,2) & between(M,3,5))

for(i in NBS$Index){
  RFM[i,10] <- "New Big Spender"
}
```

One Time Big Spender
```{r}
OTBS<-RFM %>% filter(between(R,1,2) & between(F,1,2) & between(M,3,5))

for(i in OTBS$Index){
  RFM[i,10] <- "One Time Big Spender"
}
```

```{r}
#Remove NAs
RFM_segments <- na.omit(RFM)
```

```{r}
seg_counts<- RFM_segments %>% group_by(segment) %>% summarise(count=n()) %>% mutate(proportions = round((count/sum(count)*100),2))
```

```{r}
seg_counts %>% ggplot(aes(x=segment, y=proportions, fill=segment)) + geom_bar(stat="identity") + coord_flip() + theme_bw() + ylab("Proportions (%)") + xlab("Segments")
```

## Get Response Rates

This section will focus on those to whom catalogs were sent and build on the required_catalog_extract table.
```{r}
required_catalog_extract_2 <- required_catalog_extract %>% filter(year== 2007)

required_catalog_extract_2$Recency <- current_date - required_catalog_extract_2$orderdate
```

```{r}
RFM_catalog <- required_catalog_extract_2 %>% group_by(cust_id) %>% summarize(Recency = min(Recency))
```

```{r}
cat_mail<- catalog_only %>% filter(year!=2005)
cat_mail<- cat_mail %>% group_by(cust_id) %>%  summarise(Mailed=n())
RFM_catalog <- inner_join(RFM_catalog, cat_mail, by="cust_id")
```

To get order Frequency and Average Spend, need to go back to lines table:
```{r}
required_catalog_CID <- unique(required_catalog_extract_2$cust_id)
required_catalog_ordernum <- unique(required_catalog_extract_2$ordernum)

catalog_lines <- lines %>% filter(cust_id %in% required_catalog_CID)
catalog_lines <- catalog_lines %>% filter( ordernum %in% required_catalog_ordernum)

catalog_lines_summary <- catalog_lines %>% group_by(cust_id)  %>% summarise(Avg_spend=mean(linedollars, na.rm = TRUE))

single_orders <- catalog_lines %>% group_by(cust_id,ordernum, orderdate) %>% summarise(count=n()) %>% group_by(cust_id) %>% summarise(Frequency=n())

```

```{r}
RFM_catalog <- inner_join(RFM_catalog,catalog_lines_summary, by="cust_id" )
RFM_catalog <- inner_join(RFM_catalog,single_orders, by="cust_id" )
```

Add ranks:
```{r}
RFM_catalog$R <- ntile(-RFM_catalog$Recency,5)
RFM_catalog$F <- ntile(RFM_catalog$Frequency,5)
RFM_catalog$M <- ntile(RFM_catalog$Avg_spend.x,5)
```

Combine for RFM Score
```{r}
RFM_catalog <- RFM_catalog %>% mutate(RFM = paste0(R, F, M))
```

Summarise by RFM Score:
```{r}
RFM_catalog_summarised<- RFM_catalog %>% group_by(RFM) %>% summarise(Responses=sum(Frequency),Mailed = sum(Mailed))

RFM_catalog_summarised$response_rate<- round((RFM_catalog_summarised$Responses/RFM_catalog_summarised$Mailed)*100,2)

RFM_catalog_summarised <- RFM_catalog_summarised %>% arrange(desc(response_rate))

```

RFM Economics
```{r}
cost_per_mail <- 1
unit_profit <- 30

RFM_catalog_summarised$profit <- RFM_catalog_summarised$response_rate * RFM_catalog_summarised$Responses * unit_profit
RFM_catalog_summarised$cost <- cost_per_mail * RFM_catalog_summarised$Mailed
RFM_catalog_summarised$cumprofit<- cumsum(RFM_catalog_summarised$profit)
RFM_catalog_summarised$cumcost<- cumsum(RFM_catalog_summarised$cost)
RFM_catalog_summarised$TotalMailed<-cumsum(RFM_catalog_summarised$Mailed)

RFM_catalog_summarised$ROI <- RFM_catalog_summarised$cumprofit/RFM_catalog_summarised$cost
```

plot:
```{r}
RFM_catalog_summarised %>% ggplot(aes(y=ROI,x=TotalMailed)) +geom_line()
```

# What effect do frequency catalog mailing and emailing have on line dollars?
Begin with customers mailed and sent a catalog:
```{r}
received_both <- intersect(email_contact$cust_id, catalog_contact$cust_id)
both_mailed<- contacts %>% filter(cust_id %in% received_both) %>% group_by(cust_id, contacttype) %>% summarise(count=n())

both_mailed <- both_mailed %>% spread(key=contacttype, value= count)
```
Distribution of email and catalog sent:
```{r}
ggplot() +geom_density(data=both_mailed, aes(x=both_mailed$E, col="blue")) + geom_density(data=both_mailed, aes(x=both_mailed$C, col="red")) + scale_color_identity(name= "Channel",labels=c("Email", "Catalog Mail"),guide = "legend")  + xlab("Quantity Mailed") + theme_bw()
```
 Correlation between Mail sent and Catalog sent:
 
```{r}
ggplot() +geom_point(data=both_mailed, aes(x=both_mailed$E, y=both_mailed$C, alpha=0.4)) + xlab("Quantity Emailed") + ylab("Quantity Catalog Mailed")+ theme_bw()
```
```{r}
both_lines <- lines %>% filter(orderdate>="2005-01-06" & orderdate<="2007-12-26")
both_lines <- lines %>% group_by(cust_id) %>%  summarise(Total_spend=sum(linedollars))
both_mailed <- inner_join(both_mailed, both_orders, by="cust_id")
```

Run linear regression:
```{r}
mod1<- lm(Total_spend ~ C + E, data = both_mailed)
summary(mod1)
```
```{r}
mod2<- lm(Total_spend ~ C*E, data = both_mailed)
summary(mod2)
```

# CLV Calculation
Find customers acquired in 2000:
```{r}
pre_2001 <- summary %>% select(c("Cust_ID","AcqDate"))
pre_2001 <- pre_2001 %>% filter(between(as.numeric(AcqDate),200000,200012))
```

```{r}
CLV<- orders_new %>% group_by(cust_id.x,year) %>% summarise(revenue = sum(linedollars))
CLV<-as.data.frame(CLV)
CLV <- CLV %>% spread(key="year", value="revenue")
```

Filter for customers acquired in 2000:
```{r}
CLV_2000 <- CLV %>% filter(cust_id.x %in% pre_2001$Cust_ID)
CLV_2000[is.na(CLV_2000)]<-0 #Fill NAs
```
